---
title: "Practical Implications of Using the Inverse Transformation in MLR"
author: "Lydia Gibson"
format: 
  revealjs:
    preview-links: true
    theme: simple
    footer: Cal State East Bay STAT 694 Final Presentation
---

# Introduction

```{r echo=FALSE, warning=FALSE, message=FALSE}
dat1 <- read.csv("women-stem.csv")

library(pacman)

suppressWarnings(p_load(dplyr, ggplot2, ggpubr, scales, MASS, car, lmtest, 
                        ggrepel, faraway, ggcorrplot, GGally, lindia, see, performance, ggstatsplot, rstantools, PMCMRplus, cowplot, parameters, report, datawizard))

options(scipen = 100) # remove scientific notation
```

- In [previous research](https://github.com/lgibson7/Gender-Wage-Inequality-in-STEM), I used **multiple linear regression (MLR)** to explore the relationship between gender demographics and median salary of STEM major categories. 
- My final model used the inverse transformation of the response variable to improve the model fit.
- While transforming response and/or explanatory variables can lead to a better fitting model, often it is not easily understood by the average person. 

# Outline

Problem

Data Source

Background

Methods

Results

Conclusion

Further Research

# Problem

How much prediction power is lost by **not** using a transformed response variable in a multiple linear regression model?

# Data Source

- The [data](https://github.com/fivethirtyeight/data/blob/master/college-majors/women-stem.csv) was obtained from the American Community Survey (ACS) 2010-2012 Public Use Microdata Series (PUMS). 

- It has 76 observations, each representing a STEM major, and 9 variables: `Rank`, `Major_code`, `Major`, `Major_category`, `Total`, `Men`, `Women`, `ShareWomen`, `Median`. 


# Background

## Major Category by Share of Women

```{r echo=FALSE, warning=FALSE, message=FALSE}

# remove Rank, Major_code, and Major
dat2 <- dat1[,-c(1,2,3)] 

# Get totals for men and women for each major category
dat_stats <- rbind(
  
#   Get totals for men
  dat2 %>% group_by(Major_category) %>%
    summarize(Grand_Total = sum(Men), Proportion=Grand_Total/sum(Total)) %>%
    mutate(Sex="Men", labelpos=Proportion/2),
  
# Get totals for women
dat2 %>% group_by(Major_category) %>%
 summarize(Grand_Total = sum(Women), Proportion=Grand_Total/sum(Total)) %>%
 mutate(Sex="Women", labelpos=1 - (Proportion/2))) %>% 
mutate(Sex = Sex %>% factor(levels=c("Men", "Women")))

# bar chart
p7 <- dat_stats %>% ggplot(aes(x=Major_category, y=Proportion, fill=Sex)) +
  stat_summary(geom = "bar", position="fill") +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(x="Major Category", y="Proportion of Gender (%)") +
  scale_x_discrete(labels = scales::label_wrap(15)) +
  scale_fill_manual(values = c( "#F1F1F1", "#FCA3B5") 
                    )

p7 <- p7 + theme(legend.position = "top")
p7


```

## Median by Major Category

```{r}
# violin plot
p8 <- ggstatsplot::ggbetweenstats(
  data  = dat1,
  x     = Major_category,
  y     = Median,
  type = "robust",
  outlier.tagging = TRUE,
  outlier.label = Major,
  package = "ggsci",
  palette = "default_jco"
  
  )

p8
```




```{r echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}

# Get the outliers
outlier_pts <- dat1 %>%filter(Median > 100000 |(Median > 60000 & Major_category == "Physical Sciences"))

jitter_plt <- dat1 %>%ggplot(aes(x=Major_category,y=Median, color=Major_category, size=ShareWomen)) +
  geom_jitter(alpha = 1/2) +# make circle transparent to show overlap
  theme(axis.text.x = element_text(angle=0, vjust=0.65),
        plot.subtitle = element_text(hjust=0.5),
        legend.position = "right") +
  geom_text(data=outlier_pts, aes(label=Major, size=0.11),nudge_y=2, vjust=-1.6, hjust=0.7, color = "black") + # label outliers
  
   labs(x="Major Category", y="Median Salary ($ 1000)") +
guides(color = F, # remove legend for color
         size = guide_legend( # change size legend
           override.aes = list(alpha = 1))) +
  # Wrap text for major categories
  scale_x_discrete(labels = scales::label_wrap(15)) +
  scale_size_binned(breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)) +
   # change scale labels of Median
  scale_y_continuous(breaks=c(40000,60000,80000,100000),
                     labels=c("40","60","80","100")) 

jitter_plt + scale_colour_viridis_d(option = "B")

```


# Method

- In this project, I will compare multiple linear regression models **with** an inverse transformation dependent response variable to one **without** an inverse transformation dependent response variable.
- The **full additive model** is without any transformations while the **reduced model** uses the inverse transformed response variable and interaction terms for better fit.

# Models


## Full Additive Model Parameters

```{r echo=FALSE}
options(scipen=3)

#datawizard::standardize(dat2)

lm_full <- lm(Median~.,data=dat2)

lm_reduced <- lm((Median^(-1)) ~ Major_category + Men + ShareWomen + Men*ShareWomen,
                 data=dat2)


```


```{r, size= "tiny"}
parameters::model_parameters(lm_full)
```

## Reduced Model Parameters


```{r, size="tiny"}
parameters::model_parameters(lm_reduced)

```


## Comparison of Models


```{r echo=FALSE, warning=FALSE, message=FALSE}

#library(performance)
#library(see)

p1 <- ggpubr::gghistogram(dat2$Median) + ggtitle("Response Variable Without Transformation")

p2 <- ggpubr::gghistogram(dat2$Median^-1) + ggtitle("Response Variable With Transformation")

p3 <-lindia::gg_boxcox(lm_full) + ggtitle("Box Cox Plot for Full Additive Model")


p4 <- lindia::gg_boxcox(lm_reduced) + ggtitle("Box Cox Plot for Reduced Model")


plot_grid(p1, p2, p3, p4)
```




## Diagnostic Plots for Full Additive Model

```{r}
#lindia::gg_diagnose(lm_full)
performance::check_model(lm_full)
```

## Diagnostic Plots for Reduced Model

```{r}
#lindia::gg_diagnose(lm_reduced)
performance::check_model(lm_reduced)
```



# Results

## Metrics Comparison

```{r}
performance::compare_performance(lm_reduced, lm_full, metrics = "common")
```

## Model Prediction

*statistical output*

```{r eval=FALSE}
report::report(lm_reduced)
report::report(lm_full)
```





# Conclusion

- Bullet point 1
- Bullet point 2
- Bullet point 3

# Further Research

-   I would like to use the ACS Application Programming Interface (API) to export more recent data and run the multiple linear regression models using the [TidyModels](https://github.com/tidymodels) framework.

-   I would like to compare the base R statistical data visualizations to those from the various ggplot2 extension packages ([`ggpubr`](https://github.com/kassambara/ggpubr), [`see`](https://github.com/easystats/see), [`performance`](https://github.com/easystats/performance), [`lindia`](https://github.com/yeukyul/lindia), [`ggstatsplot`](https://github.com/IndrajeetPatil/ggstatsplot)) used in this presentation.

# Acknowledgements

- I would like to thank my colleagues, Sara Hatter and Ken Vu, with whom I collaborated on the previous research project [*Gender Wage Inequality in STEM*](https://github.com/lgibson7/Gender-Wage-Inequality-in-STEM).

- I would like to thank the FiveThirtyEight blog for uploading the data behind their story, [*The Economic Guide To Picking A College Major*](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/), which was used for this analysis.

- I would like to thank Prof. Eric Suess for his assistance with this research.

# Appendix

This presentation can be viewed at: <https://lgibson7.quarto.pub/women-in-stem>. 

Code for this presentation can be found at: <https://github.com/lgibson7/Women-in-STEM>.



